---
title: "Introduction to Linear Models"
format:
  html:
    code-fold: true
jupyter: python3
---

```{python}
import os
os.chdir("..")
```

```{python}
import numpy as np
```



# Introduction to Linear Algebra

- We will use $n$ to denote the number of observations or units in our sample and use $p$ 
  to denote the number of variables measured on each unit.
- In this courses, bold uppercase letters are typically used to denote matrices, and bold 
  lowercase letters are typically used to denote vectors.
- R and Python are two of the most widely used programming languages in the field of data 
  science, machine learning, and statistical computing.
- Gained immense popularity due to their versatility, extensive livraries, and active community 
  support.
- Many researchers and practitioners now adopt a hyprid approach leveragiing the strengths of 
  both languages.

## Python 

- General- purpose language that's good for a variety of use cases, including machine learning, web development, 
  and large-scale application
- Python is considered easy to learn, especially for those wiht experience in Java or C++.
- You ccan install Anaconda, which is a package manager, environment manager, and Python distribution with many 
  third party FOSS packages.

## R 

- Primarily used for statistical analysis and  data visualization.
- R is well-suited for data scientists and researhers who want to explore and present data.
- It has powerful packages for data analysis and visualization

# Vactor and Matrix 

- Vector 
  $$\mathbf(x)_{p \times 1} = \begin{bmatrix} x_1 \\ x_2 \\ \vdots \\ x_p \end{bmatrix}$$
- Matrix 
  $$\mathbf{X}_{n \times p} = \begin{bmatrix} x_{11} & x_{12} & \cdots & x_{1p} \\ x_{21} & x_{22} & \cdots & x_{2p} \\ \vdots & \vdots & \ddots & \vdots \\ x_{n1} & x_{n2} & \cdots & x_{np} \end{bmatrix}$$

## Some properties about matrices

- A square matrix is symmetric if $A = A^T$
- If a square matrix $A$ has elements $a_{ij}$, then $A$ is symetric if $a_{ij} = a_{ji}$. 
- The transpose of the column vector $x_{p \times 1}$ is row vector.
  $$x^T = \begin{bmatrix} x_1 & x_2 & \cdots & x_p \end{bmatrix}$$

```{r}
set.seed(787)
A <- matrix(rnorm(n=25), nrow=5, ncol=5)
A
```

```{r}
np.random.seed(787)
A = np.random.normal(size=(5,5))
A
```

## Basic Concepts of Matrix Alebra 

- Two vectors can be added if they have the same dimension.
$$ \mathbf{X} + \mathbf{Y} = \begin{bmatrix} x_1 \\ x_2 \\ \vdots \\ x_p \end{bmatrix} + \begin{bmatrix} y_1 \\ y_2 \\ \vdots \\ y_p \end{bmatrix} = \begin{bmatrix} x_1 + y_1 \\ x_2 + y_2 \\ \vdots \\ x_p + y_p \end{bmatrix}$$pip install pynvim jupyter_client cairosvg plotly kaleido pnglatex pyperclip

- A vector can be added if they have the same dimension. 
  $$c \cdots \mathbf{X} = \begin{bmatrix} c \cdot x_1 \\ c \cdot x_2 \\ \vdots \\ c \cdot x_p \end{bmatrix}$$

## Examples: Vector Operations 

1. Transpose:
$$\mathbf{X} = \begin{bmatrix} 2 \\ 1 \\ -4 \end{bmatrix = \begin{bmatrix} 2 & 1 & -4 \end{bmatrix}$$

```{python}
x = np.array([[1,2,3], [4,0,4], [1,0,1]])
np.linalg.matrix_rank(x)

```

$GAG = G$


