---
title: "Measure of linear associations"
format:
  html:
    code-fold: true
---

## Introduction

- We are interested to see if from a observed sample some of the 
  variables of interest are associated.
- Measure the strength of the relationship among the sample

## Bivariate Normal Distribution

A random variable $X= (X_1,X_2)' \sim N_2 (\mu, \Sigma)$

$$
f9x|\mu, \Sigma) = \frac{1}{(2\pi)|\Sigma|^\frac{1}{2}}
\exp \big{\{} -\frac{1}{2}(x - \mu)'\Sigma^{-1}(x-\mu)
$$

## Hypothesis testing for the correlation coefficient $\rho$

Let $(X_1, Y_1), \ldots, (X_n, Y_n)$ be a random sample from a bivariate normal distribution.

- Consider the following Hypothesis
$$
H_0: \rho = 0 \quad \text{against} \quad H_1: \rho \neq 0
$$

- Test statistics

| Pearson's correlation | Spearman rank | Kendall's $\tau$ |
|-----------------------|---------------|------------------|
| $t = \frac{r\sqrt{n-2}}{\sqrt{1-r^2}}$ | $t = r_s\sqrt{n-2}$ | $z^* = \frac{3\sqrt{n(n-1)}}{2} \tau$ |

- Pearson correlation is also the fisher transormation $\log \frac{}{}
- Pearson: $t^*$

## Example: Blood pressure measurements 

Pincherle and Robinson (1974) note a marked interobserver variation in
blood pressure readings. They found that doctors who read high on
systolic tended to read high on diastolic. Data contains the means systolic
and diastolic blood pressure readings by 14 doctors. We wish to compute
a measure of the strength of the relationship. Under the assumptions that
these 14 doctors constitute a random sample from a population of doctors,
we wish to known whether we may conclude from the data that there is a
direct relationship between systolic and diastolic readings
Example: Blood pressure measurements

| systolic | diastolic |
|----------|-----------|
| 142      | 90        |
| 140      | 74        |
| 132      | 84        |
| 132      | 78        |
| 136      | 86        |
| 141      | 86        |
| 144      | 89        |
| 140      | 89        |
| 141      | 88        |
| 132      | 82        |
| 131      | 85        |
| 136      | 84        |
| 144      | 86        |
| 133      | 86        |


### R: Pearson Correlation

```{r}
systolic <- c(141.8,140.2,131.8,132.5,135.7,141.2,143.9,
140.2,140.8,131.7,130.8,135.6,143.6,133.2)
diastolic <- c(89.7,74.4,83.5,77.8,85.5,86.5,89.4,
89.3,88.0,82.2,84.6,84.4,86.3,85.9)
## cor(systolic,diastolic,method="pearson")
cor.test(systolic,diastolic,method = "pearson")
```

### R: Spearman Rank Correlation

```{r}
pressure <- cbind(systolic,diastolic)
ranks <- apply(pressure,2,rank)
d <- apply(ranks,1,diff); n <- nrow(pressure)
rs <- 1 - 6* sum(d^2)/(n*(n^2-1))
c("rs"=rs,"p-value" = pnorm(rs*sqrt(n-1),lower.tail=FALSE))
```
- Note that the Spearman is significant but pearson. This might significate that 
  the relationship is not linear.
```{r}
cor.test(systolic,diastolic,method = "spearman")
```
- The p-value in the previous e output shows the "true" p-value 


### R: Kendall's $\tau$

```{r}
rk <- cor(systolic,diastolic,method = "kendall")
c("tau" = rk,"p-value" =pnorm(3*rk*sqrt(n*(n-1)/((2*(2*n+5)))),lower.tail=FALSE))
```

```{r}
cor.test(systolic,diastolic,method="kendall")
```


