\documentclass[10pt, oneside]{article}
\usepackage{amsmath, amsthm, amssymb, calrsfs, wasysym, verbatim, bbm, color, graphics, geometry}

\geometry{tmargin=.75in, bmargin=.75in, lmargin=.75in, rmargin = .75in}

\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\Cdot}{\boldsymbol{\cdot}}

\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
\newtheorem{conv}{Convention}
\newtheorem{rem}{Remark}
\newtheorem{lem}{Lemma}
\newtheorem{cor}{Corollary}


\title{ESMA 6205: Introduction Linear Models}
\author{Alejandro Ouslan}
\date{Spring 2025}

\begin{document}

\maketitle
\tableofcontents

\vspace{.25in}

\section{Introduction to the Scientific Method}
Process by which scientist, collectively and over time, endeavor to construct an accurate representation of
the world.
\begin{itemize}
	\item Observation and description of a phenomenom or group of phenomena.
	\item Formulation of an hypothesis to explain the phenomena.
	\item Use of the hypothesis to predict the existence of other phenomna, or to predict quantitatively the results of new observation
	\item Performance of experimental tests of the predictions by several independent experimenters and properly performed experiments.
\end{itemize}

\section{What is statistics?}

\begin{itemize}
	\item Statistics science of learning from data, and of measuring, controlling, and communication uncertainty.
	\item Provides the navigation essential for controlling, and societal advances
\end{itemize}

\section{Why become an statistician/data sciences?}
\begin{itemize}
	\item "The best thing about being a statistician is that you get to play in everyone else's backyard",
	      John Tukey, Princeton University.
	\item Data science combines math and statistics, specialized programing, advanced analytics, artificial intelligence (AI)
	      and machine learning with specific subject matter expertise to uncover insights in data.
	\item Statiscian apply statistical thinking and methods to a wide variety of scientific, social, and buisness
	      endeavors in such areas as astronomy, biology, education, economics, engineering, genetics, marketing , medicine,
	      psychology, public health sports, among many.
\end{itemize}

\section{Paths in Stats}
Like many choices in life, frequentist and bayesian.
\begin{itemize}
	\item \textbf{Statistics Theory:} provides an underlyin rational and provides a consistent basis for the choice of methodology.
	\item \textbf{Applied Statistics:} comprises descriptive statisics and the application of inferential statistics.
\end{itemize}

\section{Notation Notation and Simple Matrix Algebra}

\begin{itemize}
	\item \textit{n}: number of distinct data points, or observvations, in our sample.
	\item \textit{p}: number of distinct that are available for use.
	\item \textbf{X}: matrix
	\item \textbf{y}: Vector
\end{itemize}

\section{Why linear models?}
\begin{itemize}
	\item Linear models are a fundamentals in statistics, data science and machine learning to understand relationships
	      between variables.
	\item Linear
\end{itemize}item


\end{document}
